### 2018-04-24 ident/docs/draft_3.tex
complete section 3.1

	\subsection{Structural and practical identifiability of parameters in kinetic models}	
	In the Introduction, we briefly metioned that the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}. Below, we provide a formal definition of structural and practical identifiability of parameters.

@@

	\section{Results:}
	First, in Section \ref{sec:example}, we demonstrate the use of the methodology that we described in section 2.1 to practically identify parameters in flux v1 of the small gluconeogenic network (Figure \ref{fig:network}) model given in Section \ref{sec:small-model}. In Section 3.2 that follows, we show how the demonstrated methodology is capable of identifying and estimating parameters for fluxes $v_1$, $v_2$ and $v_5$ using steady state flux values and metabolite concentrations. We discuss the ability of the proposed methodology to determine the structural and practical non-identifiability of parameters modeling $v_3$ in Section \ref{sec:proof}. The contribution of the uncertainty in the data (due to differences between in vivo and in vitro kinetics) towards identifying parameters in enzyme kinetic models is discussed in Section \ref{sec:uncertainty}. The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:design} that follows.

@@

	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b). These expressions have the form shown in Equation (\ref{eq:theta-eq}).
	\begin{subequations}\label{eq:v1_par}
		\begin{align}		
		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
		\end{align}
	\end{subequations}
	In Equation (\ref{eq:v1_par}), the denominator of the right hand side expression is used to test the identifiability of parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b) for different available experimental data combinations. Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b).

	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is also available, in addition to the measured concentration of the metabolites and fluxes. In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$. Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) above is substituted with $V_1^{max} = k_1^{cat}E$ instead.

	\subsection{Section Title}\label{sec:initial_analysis}
	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=.8\textwidth,height=.6\textheight,keepaspectratio]{figures/figure6/v1_k1cat_v2_v5_2_v3_ck_1_parameter_values}}
		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, c) $v_5$, and d)$v_3$. For $v_1$, we have assumed that enzyme concentration is available and have accordingly identified and estimated $k_1^{cat}$, as opposed to $V_1^{max}$. The parameter values for only the second root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(2)$) is shown, since $K_e^{fdp}(1)$ is not estimated by any combination of two experiments, and $V_e^{max}$ is estimated by all combinations. Only one of the two roots for $v_3$ is shown in panel d. The estimated data for the second root has a similar distribution to that of the first root and is shown in the Supplementary Information. Data is generated using the Convenience Kinetic model for allosteric regulation for $v_3$.}\label{fig:ident_values}
	\end{figure}
	\subsection{Relationship between structurally and practically parameter identifiability}\label{sec:proof}
	This section will deal with structural and practical identifiability of parameters in v1, v2, v3 and v5. This section will only focus on the identifiability and not issues related to experimental design. Focus will be on v3 and v5 (which have non-identifiability issues). Also, we could use a better name for this section.	
	\subsection{Parameter non-identifiability due to uncertainty in Experimental Data}\label{sec:uncertainty}
	This will cover issues related to the model used for v3. Convenience kientic model vs MWC model for data generation. Should this uncertainty also include noise?			
	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
	Final section talking about distribution of experiments to obtain informative data.

>>>>

	\subsection{Structural and practical identifiability of parameters in kinetic models}\label{sec:ident_def}
	In the Introduction, we briefly metioned that the ability to estimate unique parameter values from available experimental data is governed by the identifiability of these parameters in the model \parencite{Ljung1994,Vanlier2012,Berthoumieux2013,Raue2014}. Below, we provide a formal definition of structural and practical identifiability of parameters.

@@

	\section{Results}
	First, in Section \ref{sec:example}, we demonstrate the use of the methodology that we described in Section 2.1 to practically identify parameters in flux $v_1$ of the small gluconeogenic network (Figure \ref{fig:network}) model given in Section \ref{sec:small-model}. We discuss the ability of the proposed methodology to determine the structural and practical non-identifiability of parameters modeling $v_1$, $v_3$ and $v_5$ in Section \ref{sec:proof}. 	
	In Section \ref{sec:initial_analysis} that follows, we show how the demonstrated methodology is capable of identifying and estimating parameters for fluxes $v_1$, $v_2$, $v_3$ and $v_5$ using steady state flux values and metabolite concentrations. 	
	The contribution of the uncertainty in the data (due to differences between in vivo and in vitro kinetics) towards identifying parameters in enzyme kinetic models is discussed in Section \ref{sec:uncertainty}. The various ways in which this information can be used for designing experiments to generate data that can facilitate estimation of identifiable parameters are discussed in Section \ref{sec:design} that follows.

@@

	Solving this simultaneous system of equations in $\mathbb{R}^2$ using Mathematica (Wolfram Research, USA), a computer algebra system, we get $p=2$ nonlinear algebraic expressions for parameters $V_1^{max}$ (Equation \ref{eq:v1_par}a) and $K_1^{ac}(ne)$ (Equation \ref{eq:v1_par}b). These expressions have the form shown in Equation (\ref{eq:theta-eq}).
	\begin{subequations}\label{eq:v1_par}
		\begin{align}		
		\theta_1 &= V_1^{max} = \frac{v_1^{(1)}v_1^{(2)}(ac^{(1)}-ac^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}\\
		\theta_2 &= K_1^{ac}(ne) = \frac{ac^{(1)}ac^{(2)}(v_1^{(1)}-v_1^{(2)})}{v_1^{(2)}ac^{(1)}-v_1^{(1)}ac^{(2)}}
		\end{align}
	\end{subequations}
	To test the practical identifiability of the parameters in Equation \ref{eq:v1_par}, we substitute any suitable in silico experimental data and determine the value of the denominator of the right hand side expression. Since the enzyme binding constant ($K_1^{ac}(ne)$) and the maximum reaction rate ($V_1^{max}$) cannot be negative, we can further constrain the criteria for identifiability for both these parameters by saying that the evaluated expressions in Equation (\ref{eq:v1_par}) should be non-negative (Figure \ref{fig:ident-flowchart}b). The parameter values that are obtained for $V_1^{max}$ and $K_1^{ac} (ne)$ by substituting in silico steady state experimental data are shown in Supplementary Figure \ref{fig:v1_v1max_ck_values}.

	We can also apply the proposed methodology to practically identify parameters in $v_1$ under the assumption that the protein concentration for the enzyme \textit{E} is also available, in addition to the measured metabolite concentrations and fluxes. In doing so, we get two expressions similar to the one shown in Equation (\ref{eq:v1_par}) for $k_1^{cat}$ and $K_1^{ac}$. Here, the value of $V_1^{max}$ in Equation (\ref{eq:flux1a}) is substituted with $V_1^{max} = k_1^{cat}E$ instead. The corresponding identifiability expressions for $k_1^{cat}$ and $K_1^{ac}$ are given in Equation (\ref{eq:v1_k1cat}).	
	\begin{subequations}\label{eq:v1_k1cat}
		\begin{align}
		k_1^{cat} = \frac{ v_1^{(1)}v_1^{(2)} \left( ac^{(1)}- ac^{(2)}\right)}{v_1^{(2)}ac^{(1)}E^{(1)} - v_1^{(1)}ac^{(2)}E^{(2)}}\\
		K_1^{ac} = \frac{ac^{(1)}ac^{(2)}\left(v_1^{(1)}E^{(2)} - v_1^{(2)}E^{(1)}\right)}{v_1^{(2)}ac^{(1)}E^{(1)} - v_1^{(1)}ac^{(2)}E^{(2)}}
		\end{align}
	\end{subequations}	
	We show the parameter value for $k_1^{cat}$ and $K_1^{ac}$ that are obtained through the practical identifiability analysis in Figure \ref{fig:ident_values}a when in silico experimental data is substituted in Equation (\ref{eq:v1_k1cat}). Through Equation (\ref{eq:v1_k1cat}) and Figure \ref{fig:ident_values}a we are able to show that the uncertainty in the experimental flux values can be resolved through the incorporation of the available enzyme concentrations to get good estimations for the enzyme kinetic parameters. Thus, having more experimental information can help resolve practical identifiability.

	\begin{figure}[!tbhp]
		\centering{\includegraphics[width=.8\textwidth,height=.6\textheight,keepaspectratio]{figures/figure6/v1_k1cat_v2_v5_2_v3_ck_1_parameter_values}}
		\caption{Distribution of predicted parameter values when performing practical identifiability analysis using closed-form solutions for each parameter in flux a) $v_1$, b) $v_2$, c) $v_5$, and d)$v_3$. For $v_1$, we have assumed that enzyme concentration is available and have accordingly identified and estimated $k_1^{cat}$, as opposed to $V_1^{max}$. The parameter values for only the second root of $K_e^{fdp}$ in $v_5$ ($K_e^{fdp}(2)$) is shown, since $K_e^{fdp}(1)$ is not estimated by any combination of two experiments, and $V_e^{max}$ is estimated by all combinations. Only one of the two roots for $v_3$ is shown in panel d. The estimated data for the second root has a similar distribution to that of the first root and is shown in the Supplementary Information. Data is generated using the Convenience Kinetic model for allosteric regulation for $v_3$.}\label{fig:ident_values}
	\end{figure}

	In the following section we present results from the identifiability analysis of fluxes $v_2$, $v_3$ and $v_5$ in the small metabolic network (Figure \ref{fig:network}), using the methodology (Figure \ref{fig:ident-flowchart}) that we have demonstrated above for $v_1$.

	\subsection{Relationship between structural and practical parameter identifiability}\label{sec:proof}
	As mentioned in Section \ref{sec:ident_def}, by definition, it should be possible to estimate unique parameter values based on the model structure and available experimental data for any structurally and practically identifiable parameter. For the proposed methodology (Figure \ref{fig:ident-flowchart}) to work, it should be possible to obtain closed form solutions for each parameter in the enzyme kinetic model for each flux as shown in Equation (\ref{eq:theta-eq}). Since the ability to obtain closed-form solutions for each parameter is dependent on the model structure, any parameter that has non-unique closed-form solutions can be called a structurally non-identifiable parameter.  However, if the number of solutions that the parameter has are finite, then the parameter is only locally structurally identifiable.

	We demonstrated the identifiability of parameters modeling $v_1$ in Section \ref{sec:example}. We have shown that the parameters have only one unique closed-form solution, and accordingly are structurally identifiable. Since $v_2$ is also expressed using the Michaelis-Menten model, just like $v_1$, we find that the parameters ($V_2^{max}$ and $K_2^{pep}$) are also structurally identifiable. The closed-form expressions for these parameters are similar to the ones shown in Equation \ref{eq:v1_par}, with \textit{ac} replaced by \textit{pep}, and $v_1$ replaced by $v_2$.

	As shown in Figure \ref{fig:ident-flowchart} and illustrated for $v_1$ in Section \ref{sec:example}, to determine the practical identifiability of parameters we test for the existence of a non-zero denominator of the closed-form expressions of the parameters. Additionally, we also check for the physiological feasibility of the parameter values (Figure \ref{fig:ident-flowchart}b) that are obtained through this analysis. If the resulting parameter values obtained from unique closed-form expressions and various combinations of experimental data are also unique, then the parameter is both structurally and practically identifiable. Accordingly, both $v_1$ and $v_2$ are not only structurally identifiable due to the presence of unique closed-form expressions for their parameters, they are also practically identifiable because the parameters in the respective models possess unique values based on distinct combinations of experimental data (Figure \ref{fig:ident_values}a and b).

	However, we find that the parameters used to model $v_3$ using the Convenience kinetics rate law, and $v_5$ using the Hill kinetic rate law are not structural identifiable under certain conditions. First, for $v_3$, we find that the parameters $V_3^{max}$, $K_3^{fdp}$ and $K_3^{pep}$ have two different closed-form solutions. Thus, based on the presence of non-unique but finite number of possible solutions for these parameters we can classify $v_3$ as a locally structurally identifiable flux.

	We would like to point out that we used the Convenience kinetics rate law model for $v_3$ due to the inability of the computer algebra system to generate closed-form expressions for parameters in the Monod-Wyman-Chageaux model (Section \ref{sec:small-model}). However, we find that the Convenience kinetics model is not structurally identifiable in its current form. In order to alleviate this issue, we reduced the dimension of the parameter space for $v_3$. Originally, $\theta \in \mathbb{R}^3$ for $v_3$. By reducing the dimension of $\theta$ to $\mathbb{R}^2$, we were able to obtain a structurally identifiable model for $v_3$. To reduce the dimension of the parameter space for $v_3$, we fix either $K_3^{fdp}$ or $K_3^{pep}$ as a known quantity, and identify the other unfixed parameter along with $V_3^{max}$. This results in unique closed-form expressions for both $V_3^{max}$ and the other unfixed parameter ($K_3^{pep}$ or $K_3^{fdp}$).

	$v_5$ describes a transcription/translation reaction using Hill kinetics. We apply our proposed methodology to identify parameters modeling $v_5$ using only the available experimental data on the metabolite concentrations and the fluxes within the metabolic network. We could not obtain closed form solutions for parameters $V_e^{max}$, $K_e^{fdp}$ and $n_e$ in $v_5$ using the computer algebra system. So, as illustrated earlier for $v_3$, we resorted to reducing the dimension of the parameter space by fixing one of the three parameters, the Hill coefficient $n_e$. With a fixed and known $n_e$, $K_e^{fdp}$ has two possible closed-form solutions, making it a locally structurally identifiable parameter. On the other hand, $V_e^{max}$ has only one unique closed-form solution, and therefore is structurally identifiable.

	This section will deal with structural and practical identifiability of parameters in v1, v2, v3 and v5. This section will only focus on the identifiability and not issues related to experimental design. Focus will be on v3 and v5 (which have non-identifiability issues). Also, we could use a better name for this section.

	\subsection{Section Title}\label{sec:initial_analysis}

	\subsection{Parameter non-identifiability due to uncertainty in Experimental Data}\label{sec:uncertainty}
	This will cover issues related to the model used for v3. Convenience kientic model vs MWC model for data generation. Should this uncertainty also include noise?			
	\subsection{A priori experimental design through practical parameter identification}\label{sec:design}	
	Final section talking about distribution of experiments to obtain informative data.
